---
title: "Coursera Course 5 Project 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading and tranforming the data
```{r loading the data}
    setwd("C:\\Users\\shubh\\Documents\\Coursera\\Reproducible Research\\Week 2")
    activity_data <- read.csv("activity.csv", header = TRUE, stringsAsFactors = FALSE)
    activity_data$date <- as.Date(activity_data$date)
    #activity_data <- activity_data[ with (activity_data, { !(is.na(steps)) } ), ]
    activity_data_no_na <- activity_data[!is.na(activity_data$steps),]
```

##Mean total number of steps
```{r mean steps per day}
    total_steps_a_day <- unname(tapply(activity_data_no_na$steps, activity_data_no_na$date, sum))
    hist(total_steps_a_day, main = "Steps per day", xlab = "Steps", ylab = "Number of Days")
    mean(total_steps_a_day, na.rm = TRUE)
    median(total_steps_a_day, na.rm = TRUE)
    
```

**Mean: 10766.19**  
**Median: 10765**

##Average number of steps for all days
```{r average daily activity}
    steps_by_interval <- aggregate(steps ~ interval, activity_data, mean)

    # create a time series plot 
    plot(steps_by_interval$interval, steps_by_interval$steps, type='l', 
        main="Average steps for all days", xlab="Interval", 
        ylab="Average steps")
```

##Missing Values  
**replaced the NA values with mean of 5 minute intervals**  
```{r missing values}
    df <- data.frame(interval = unique(activity_data$interval), steps = tapply(activity_data_no_na$steps, activity_data_no_na$interval, mean))
replace_na <- activity_data
for(i in 1:nrow(activity_data)) {
    if (is.na(activity_data$steps[i])) {
        
        replace_na$steps[i] <- df$steps[which(df$interval==activity_data$interval[i])]
    }
}
head(replace_na)
    total_steps_a_day_replaced <- unname(tapply(replace_na$steps, replace_na$date, sum))
    hist(total_steps_a_day, main = "Steps per day with NA replaced Data", xlab = "Steps", ylab = "Number of Days")
    mean(total_steps_a_day_replaced)
    median(total_steps_a_day_replaced)

```

**Old Mean: 10766.19**  
**New Mean: 10766.19**  
**Old Median: 10765**  
**New Median: 10766.19**  
**The mean did not change but the median did, the median is a little higher with the replacement of NA values**  

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays vs weekends}
    library(chron)
    replace_na$weekend <- is.weekend(replace_na$date)
    replace_na$weekend[replace_na$weekend] <- "weekend"
    replace_na$weekend[replace_na$weekend != "weekend"] <- "weekday"
    replace_na$weekend <- as.factor(replace_na$weekend)
    average_steps <- aggregate(steps ~ interval + weekend, replace_na, mean)
    library(ggplot2)
    qplot(data = average_steps, x=interval, y = steps, geom = "line", facets = weekend ~ .)

    
```

